IMPC Mouse data - Variance in sex differences

### 2019-11-06 Prep & old/prep and old- scripts/2019_Aug30_IMPC_variance_and_sex_difference.Rmd
Update Nov

meta.plot2.all.b <- gather(meta.plot2.all, trait, value, c(lnCVR,  lnRR)) #lnVR,

meta.plot2.all.b$trait <- factor(meta.plot2.all.b$trait, levels =c("lnCVR","lnRR") )#"lnVR",

@@

meta.plot2.sig.b <- meta.plot2.sig[, c("lnCVR",  "lnRR", "lnCVRsig", "lnVRsig", "lnRRsig", "GroupingTerm")]   #"lnVR",

@@

meta.plot2.sig.c$trait <- factor(meta.plot2.sig.c$trait, levels =c("lnCVR","lnRR") ) #"lnVR",

@@

Prepare data for traits with effect size ratios > 10% larger in males
```{r}
meta.plot2.over10 <- meta_clean %>% select(lnCVR,  lnRR, GroupingTerm) %>% arrange(GroupingTerm) #lnVR,

meta.plot2.over10.b <- gather(meta.plot2.over10, trait, value, c(lnCVR,  lnRR)) #lnVR,

meta.plot2.over10.b$trait <- factor(meta.plot2.over10.b$trait, levels =c("lnCVR","lnRR") ) #"lnVR",

@@

Fig2 <- ggarrange(malebias_Fig2_alltraits, malebias_Fig2_sigtraits,malebias_Fig2_over10, nrow = 3, align = "v", heights = c(1.22,1,1), labels = c("A", "B", "C"))
Fig2
ggsave("CountDataCVR_RR_only.pdf", plot = Fig2, width = 6, height = 5) 
```

###  Overall results of second order meta analysis (Figure 4a)
#### Restructure data for plotting 
Data are restructured, and grouping terms are being re-ordered
```{r}
overall3 <- gather(overall2, parameter, value, c(lnCVR,  lnRR), factor_key= TRUE) #lnVR,

lnCVR.ci <- overall3 %>% filter(parameter == "lnCVR") %>% mutate(ci.low = lnCVR_lower, ci.high = lnCVR_upper)
#lnVR.ci <- overall3 %>% filter(parameter == "lnVR") %>% mutate(ci.low = lnVR_lower, ci.high = lnVR_upper)
lnRR.ci <- overall3 %>% filter(parameter == "lnRR") %>% mutate(ci.low = lnRR_lower, ci.high = lnRR_upper)

overall4 <- bind_rows(lnCVR.ci,  lnRR.ci) %>% select(GroupingTerm, parameter, value, ci.low, ci.high) #lnVR.ci,

@@

#Significant subset for lnVR
#metacombo_male.plot3.VR <- meta.male.plot3.sig %>%
# filter(sigVR == 1) %>%
# group_by(GroupingTerm) %>%
# nest()

#metacombo_male.plot3.VR.all <- meta.male.plot3.sig %>%
 #filter(sigVR == 1) %>%
 #nest()

@@

#plot3.male.meta.VR <- metacombo_male.plot3.VR %>% 
 # mutate(model_lnVR = map(data, ~ metafor::rma.uni(yi = .x$lnVR, sei = (.x$lnVR_upper - .x$lnVR_lower)/ (2*1.96),  
#				control=list(optimizer="optim", optmethod="Nelder-Mead", maxit= 1000), verbose=F)) )

@@

#plot3.male.meta.VR.all <- metacombo_male.plot3.VR.all %>% 
#  mutate(model_lnVR = map(data, ~ metafor::rma.uni(yi = .x$lnVR, sei = (.x$lnVR_upper - .x$lnVR_lower)/ (2*1.96),  
#				control=list(optimizer="optim", optmethod="Nelder-Mead", maxit= 1000), verbose=F)) )

#plot3.male.meta.VR.all <- plot3.male.meta.VR.all %>% mutate(GroupingTerm = "All")

@@

plot3.male.meta.CVR <- bind_rows(plot3.male.meta.CVR, plot3.male.meta.CVR.all)
#plot3.male.meta.VR <- bind_rows(plot3.male.meta.VR, plot3.male.meta.VR.all)
plot3.male.meta.RR <- bind_rows(plot3.male.meta.RR, plot3.male.meta.RR.all)

@@

#plot3.male.meta.VR.b <- as.data.frame(plot3.male.meta.VR %>% group_by(GroupingTerm) %>%  
#	mutate(lnVR = map_dbl(model_lnVR, pluck(2)), lnVR_lower = map_dbl(model_lnVR, pluck(6)), 
#	lnVR_upper =map_dbl(model_lnVR, pluck(7)), lnVR_se =map_dbl(model_lnVR, pluck(3))) )[, c(1,4:7)] 
#plot3.male.meta.VR.b <- plot3.male.meta.VR.b[order(plot3.male.meta.VR.b$GroupingTerm),]

@@

#overall.male.plot3 <- full_join(plot3.male.meta.CVR.b, plot3.male.meta.VR.b) 
overall.male.plot3 <- full_join(plot3.male.meta.CVR.b, plot3.male.meta.RR.b)

@@

Restructure MALE data for plotting 
```{r}
overall3.male.sig <- gather(overall.male.plot3, parameter, value, c(lnCVR,  lnRR), factor_key= TRUE) #lnVR,

lnCVR.ci <- overall3.male.sig  %>% filter(parameter == "lnCVR") %>% mutate(ci.low = lnCVR_lower, ci.high = lnCVR_upper)
#lnVR.ci <- overall3.male.sig  %>% filter(parameter == "lnVR") %>% mutate(ci.low = lnVR_lower, ci.high = lnVR_upper)
lnRR.ci <- overall3.male.sig  %>% filter(parameter == "lnRR") %>% mutate(ci.low = lnRR_lower, ci.high = lnRR_upper)

overall4.male.sig <- bind_rows(lnCVR.ci,  lnRR.ci) %>% select(GroupingTerm, parameter, value, ci.low, ci.high) #lnVR.ci,

@@

```{r}
overall3.female.sig <- gather(overall.female.plot3, parameter, value, c(lnCVR,  lnRR), factor_key= TRUE) #lnVR,

lnCVR.ci <- overall3.female.sig  %>% filter(parameter == "lnCVR") %>% mutate(ci.low = lnCVR_lower, ci.high = lnCVR_upper)
#lnVR.ci <- overall3.female.sig  %>% filter(parameter == "lnVR") %>% mutate(ci.low = lnVR_lower, ci.high = lnVR_upper)
lnRR.ci <- overall3.female.sig  %>% filter(parameter == "lnRR") %>% mutate(ci.low = lnRR_lower, ci.high = lnRR_upper)

overall4.female.sig <- bind_rows(lnCVR.ci,  lnRR.ci) %>% select(GroupingTerm, parameter, value, ci.low, ci.high) #lnVR.ci,

@@

#overall.male.plot3.perc <- full_join(plot3.male.meta.CVR.perc.b, plot3.male.meta.VR.perc.b) 
overall.male.plot3.perc <- full_join(plot3.male.meta.CVR.perc.b, plot3.male.meta.RR.perc.b)

@@

Restructure data for plotting : Male biased, 10% difference
```{r}
overall3.perc <- gather(overall.male.plot3.perc, parameter, value, c(lnCVR, lnRR), factor_key= TRUE) #lnVR,

lnCVR.ci <- overall3.perc %>% filter(parameter == "lnCVR") %>% mutate(ci.low = lnCVR_lower, ci.high = lnCVR_upper)
#lnVR.ci <- overall3.perc  %>% filter(parameter == "lnVR") %>% mutate(ci.low = lnVR_lower, ci.high = lnVR_upper)
lnRR.ci <- overall3.perc  %>% filter(parameter == "lnRR") %>% mutate(ci.low = lnRR_lower, ci.high = lnRR_upper)

overall4.male.perc <- bind_rows(lnCVR.ci,  lnRR.ci) %>% select(GroupingTerm, parameter, value, ci.low, ci.high) #lnVR.ci,

@@

#overall.plot3.perc <- full_join(plot3.meta.CVR.perc.b, plot3.meta.VR.perc.b) 
overall.plot3.perc <- full_join(plot3.meta.CVR.perc.b, plot3.meta.RR.perc.b)

@@

Restructure data for plotting
Female bias, 10 percent difference
```{r}
overall3.perc <- gather(overall.plot3.perc, parameter, value, c(lnCVR, lnRR), factor_key= TRUE) #lnVR,

lnCVR.ci <- overall3.perc %>% filter(parameter == "lnCVR") %>% mutate(ci.low = lnCVR_lower, ci.high = lnCVR_upper)
#lnVR.ci <- overall3.perc  %>% filter(parameter == "lnVR") %>% mutate(ci.low = lnVR_lower, ci.high = lnVR_upper)
lnRR.ci <- overall3.perc  %>% filter(parameter == "lnRR") %>% mutate(ci.low = lnRR_lower, ci.high = lnRR_upper)

overall4.perc <- bind_rows(lnCVR.ci,  lnRR.ci) %>% select(GroupingTerm, parameter, value, ci.low, ci.high) #lnVR.ci,

@@

Fig3 <- ggarrange(Metameta_Fig3_alltraits, Fig3.bottom, ncol = 1, nrow = 2, heights = c(1.4, 2.5))
Fig3
#ggsave("MetaResultsCVR_RR.pdf", plot = Fig3, width = 9, height = 6) 
```

>>>>

meta.plot2.all.b <- gather(meta.plot2.all, trait, value, c(lnCVR, lnVR, lnRR))

meta.plot2.all.b$trait <- factor(meta.plot2.all.b$trait, levels =c("lnCVR","lnVR","lnRR") )

@@

meta.plot2.sig.b <- meta.plot2.sig[, c("lnCVR", "lnVR", "lnRR", "lnCVRsig", "lnVRsig", "lnRRsig", "GroupingTerm")]

@@

meta.plot2.sig.c$trait <- factor(meta.plot2.sig.c$trait, levels =c("lnCVR","lnVR","lnRR") )

@@

Prepare data for traits with effect size ratios > 10% larger in males
```{r}
meta.plot2.over10 <- meta_clean %>% select(lnCVR, lnVR, lnRR, GroupingTerm) %>% arrange(GroupingTerm)

meta.plot2.over10.b <- gather(meta.plot2.over10, trait, value, c(lnCVR, lnVR, lnRR))

meta.plot2.over10.b$trait <- factor(meta.plot2.over10.b$trait, levels =c("lnCVR","lnVR","lnRR") )

@@

Fig2 <- ggarrange(malebias_Fig2_alltraits, malebias_Fig2_sigtraits,malebias_Fig2_over10, nrow = 3, align = "v", heights = c(1.22,1,1), labels = c("A", "B", "C"))
Fig2
ggsave("Fig2.pdf", plot = Fig2, width = 6, height = 5) 
```

###  Overall results of second order meta analysis (Figure 4a)
#### Restructure data for plotting 
Data are restructured, and grouping terms are being re-ordered
```{r}
overall3 <- gather(overall2, parameter, value, c(lnCVR, lnVR, lnRR), factor_key= TRUE)

lnCVR.ci <- overall3 %>% filter(parameter == "lnCVR") %>% mutate(ci.low = lnCVR_lower, ci.high = lnCVR_upper)
lnVR.ci <- overall3 %>% filter(parameter == "lnVR") %>% mutate(ci.low = lnVR_lower, ci.high = lnVR_upper)
lnRR.ci <- overall3 %>% filter(parameter == "lnRR") %>% mutate(ci.low = lnRR_lower, ci.high = lnRR_upper)

overall4 <- bind_rows(lnCVR.ci, lnVR.ci, lnRR.ci) %>% select(GroupingTerm, parameter, value, ci.low, ci.high)

@@

#Significant subset for lnVR
metacombo_male.plot3.VR <- meta.male.plot3.sig %>%
 filter(sigVR == 1) %>%
 group_by(GroupingTerm) %>%
 nest()

metacombo_male.plot3.VR.all <- meta.male.plot3.sig %>%
 filter(sigVR == 1) %>%
 nest()

@@

plot3.male.meta.VR <- metacombo_male.plot3.VR %>% 
  mutate(model_lnVR = map(data, ~ metafor::rma.uni(yi = .x$lnVR, sei = (.x$lnVR_upper - .x$lnVR_lower)/ (2*1.96),  
				control=list(optimizer="optim", optmethod="Nelder-Mead", maxit= 1000), verbose=F)) )

@@

plot3.male.meta.VR.all <- metacombo_male.plot3.VR.all %>% 
  mutate(model_lnVR = map(data, ~ metafor::rma.uni(yi = .x$lnVR, sei = (.x$lnVR_upper - .x$lnVR_lower)/ (2*1.96),  
				control=list(optimizer="optim", optmethod="Nelder-Mead", maxit= 1000), verbose=F)) )

plot3.male.meta.VR.all <- plot3.male.meta.VR.all %>% mutate(GroupingTerm = "All")

@@

plot3.male.meta.CVR <- bind_rows(plot3.male.meta.CVR, plot3.male.meta.CVR.all)
plot3.male.meta.VR <- bind_rows(plot3.male.meta.VR, plot3.male.meta.VR.all)
plot3.male.meta.RR <- bind_rows(plot3.male.meta.RR, plot3.male.meta.RR.all)

@@

plot3.male.meta.VR.b <- as.data.frame(plot3.male.meta.VR %>% group_by(GroupingTerm) %>%  
	mutate(lnVR = map_dbl(model_lnVR, pluck(2)), lnVR_lower = map_dbl(model_lnVR, pluck(6)), 
	lnVR_upper =map_dbl(model_lnVR, pluck(7)), lnVR_se =map_dbl(model_lnVR, pluck(3))) )[, c(1,4:7)] 
plot3.male.meta.VR.b <- plot3.male.meta.VR.b[order(plot3.male.meta.VR.b$GroupingTerm),]

@@

overall.male.plot3 <- full_join(plot3.male.meta.CVR.b, plot3.male.meta.VR.b) 
overall.male.plot3 <- full_join(overall.male.plot3, plot3.male.meta.RR.b)

@@

Restructure MALE data for plotting 
```{r}
overall3.male.sig <- gather(overall.male.plot3, parameter, value, c(lnCVR, lnVR, lnRR), factor_key= TRUE)

lnCVR.ci <- overall3.male.sig  %>% filter(parameter == "lnCVR") %>% mutate(ci.low = lnCVR_lower, ci.high = lnCVR_upper)
lnVR.ci <- overall3.male.sig  %>% filter(parameter == "lnVR") %>% mutate(ci.low = lnVR_lower, ci.high = lnVR_upper)
lnRR.ci <- overall3.male.sig  %>% filter(parameter == "lnRR") %>% mutate(ci.low = lnRR_lower, ci.high = lnRR_upper)

overall4.male.sig <- bind_rows(lnCVR.ci, lnVR.ci, lnRR.ci) %>% select(GroupingTerm, parameter, value, ci.low, ci.high)

@@

```{r}
overall3.female.sig <- gather(overall.female.plot3, parameter, value, c(lnCVR, lnVR, lnRR), factor_key= TRUE)

lnCVR.ci <- overall3.female.sig  %>% filter(parameter == "lnCVR") %>% mutate(ci.low = lnCVR_lower, ci.high = lnCVR_upper)
lnVR.ci <- overall3.female.sig  %>% filter(parameter == "lnVR") %>% mutate(ci.low = lnVR_lower, ci.high = lnVR_upper)
lnRR.ci <- overall3.female.sig  %>% filter(parameter == "lnRR") %>% mutate(ci.low = lnRR_lower, ci.high = lnRR_upper)

overall4.female.sig <- bind_rows(lnCVR.ci, lnVR.ci, lnRR.ci) %>% select(GroupingTerm, parameter, value, ci.low, ci.high)

@@

overall.male.plot3.perc <- full_join(plot3.male.meta.CVR.perc.b, plot3.male.meta.VR.perc.b) 
overall.male.plot3.perc <- full_join(overall.male.plot3.perc, plot3.male.meta.RR.perc.b)

@@

Restructure data for plotting : Male biased, 10% difference
```{r}
overall3.perc <- gather(overall.male.plot3.perc, parameter, value, c(lnCVR, lnVR, lnRR), factor_key= TRUE)

lnCVR.ci <- overall3.perc %>% filter(parameter == "lnCVR") %>% mutate(ci.low = lnCVR_lower, ci.high = lnCVR_upper)
lnVR.ci <- overall3.perc  %>% filter(parameter == "lnVR") %>% mutate(ci.low = lnVR_lower, ci.high = lnVR_upper)
lnRR.ci <- overall3.perc  %>% filter(parameter == "lnRR") %>% mutate(ci.low = lnRR_lower, ci.high = lnRR_upper)

overall4.male.perc <- bind_rows(lnCVR.ci, lnVR.ci, lnRR.ci) %>% select(GroupingTerm, parameter, value, ci.low, ci.high)

@@

overall.plot3.perc <- full_join(plot3.meta.CVR.perc.b, plot3.meta.VR.perc.b) 
overall.plot3.perc <- full_join(overall.plot3.perc, plot3.meta.RR.perc.b)

@@

Restructure data for plotting
Female bias, 10 percent difference
```{r}
overall3.perc <- gather(overall.plot3.perc, parameter, value, c(lnCVR, lnVR, lnRR), factor_key= TRUE)

lnCVR.ci <- overall3.perc %>% filter(parameter == "lnCVR") %>% mutate(ci.low = lnCVR_lower, ci.high = lnCVR_upper)
lnVR.ci <- overall3.perc  %>% filter(parameter == "lnVR") %>% mutate(ci.low = lnVR_lower, ci.high = lnVR_upper)
lnRR.ci <- overall3.perc  %>% filter(parameter == "lnRR") %>% mutate(ci.low = lnRR_lower, ci.high = lnRR_upper)

overall4.perc <- bind_rows(lnCVR.ci, lnVR.ci, lnRR.ci) %>% select(GroupingTerm, parameter, value, ci.low, ci.high)

@@

Fig3 <- ggarrange(Metameta_Fig3_alltraits, Fig3.bottom, ncol = 1, nrow = 2, heights = c(1.4, 2.5))
Fig3
#ggsave("Fig3.pdf", plot = Fig3, width = 9, height = 6) 
```

### 2019-11-06 Prep & old/prep and old- scripts/2019_IMPC variance and sex difference.Rmd
Update Nov

---
title: "IMPC Mouse data - Variance in sex differences"
author: "Susanne Zajitschek,  Felix Zajitschek, Russell Bonduriansky,Robert  Brooks, Will Cornwell, Daniel Falster, Malgortaza Lagisz, Jeremy Mason, Daniel Noble, Alistair Senior & Shinichi Nakagawa"
date: "August 2019"
output:
  html_document:
    code_download: true
    code_folding: hide
    depth: 4
    number_sections: no
    theme:  flatly
    toc: yes
    toc_depth: 4
    toc_float: yes
  html_notebook:
    toc: yes
  pdf_document:
    toc: yes
    toc_depth: '4'
subtitle: Electronic Supplementary Material
---

@@

 group <- read.csv("export/ParameterGrouping.csv") 
 data <- data3
 data$parameterGroup <- group$parameter[match(data$parameter_name, group$parameter_name)] 
```

@@

```{r}
meta_clean <- meta1[ !(meta1$id %in% c(84,144,158,160,161,162,163,165,166,167,168,221,222,231)), ]
removed <-length(unique(meta_clean$parameter_name)) #218

meta_problem <- results.alltraits.grouping[ (results.alltraits.grouping$id %in% c(84,144,158,160,161,162,163,165,166,167,168,221,222,231)), ]
```

>>>>

---
title: "IMPC Mouse data - Variance in sex differences"
author: "Susanne Zajitschek,  Felix Zajitschek, Russell Bonduriansky,Robert  Brooks, Will Cornwell, Daniel Falster, Malgortaza Lagisz, Jeremy Mason, Daniel Noble, Alistair Senior & Shinichi Nakagawa"
date: "August 2019"
output:
  html_document:
    code_download: true
    code_folding: hide
    depth: 4
    number_sections: no
    theme:  cosmo
    toc: yes
    toc_depth: 4
    toc_float: yes
  html_notebook:
    toc: yes
  pdf_document:
    toc: yes
    toc_depth: '4'
subtitle: Electronic Supplementary Material
---

@@

 group <- read.csv("export/ParameterGrouping.csv") 
 data <- data3
 data$parameterGroup <- group$parameter[match(data$parameter_name, group$parameter_name)]

@@

```{r}
meta_clean <- meta1[ !(meta1$id %in% c(84,144,158,160,161,162,163,165,166,167,168,221,222,231)), ]
removed <-length(unique(meta_clean$parameter_name)) #218
```

### 2019-11-06 Prep & old/prep and old- scripts/2019_July_IMPC variance and sex difference.Rmd
Update Nov

2) Subsetting the data to choose only one datapoint per individual per trait
```{r}
# this is a necessary step for the loop across all traits 
data_subset_parameterid_individual_by_age <- function(mydata, parameter, age_min, age_center) {
  tmp <- mydata %>%
    filter(age_in_days >= age_min,
           id == parameter) %>%
    # take results for single individual closest to age_center
    mutate(age_diff = abs(age_center - age_in_days)) %>%
    group_by(biological_sample_id) %>%
    filter(age_diff == min(age_diff)) %>%
    select(-age_diff)

@@

### Heterogeneity
#### FIGURE 4 (Second-order meta analysis on heterogeneity)
#### Create matrix to store results for all traits
```{r}
results.allhetero.grouping <- as.data.frame(cbind(c(1:n), matrix(rep(0, n*30), ncol = 30))) 
names(results.allhetero.grouping) <- c("id", "sigma2_strain.CVR", "sigma2_center.CVR", "sigma2_error.CVR", "s.nlevels.strain.CVR", 
	"s.nlevels.center.CVR", "s.nlevels.error.CVR", "sigma2_strain.VR", "sigma2_center.VR", "sigma2_error.VR", "s.nlevels.strain.VR", 
	"s.nlevels.center.VR", "s.nlevels.error.VR", "sigma2_strain.RR", "sigma2_center.RR", "sigma2_error.RR", "s.nlevels.strain.RR", 
	"s.nlevels.center.RR", "s.nlevels.error.RR", "lnCVR", "lnCVR_lower", "lnCVR_upper", "lnCVR_se", "lnVR", "lnVR_lower", "lnVR_upper", 
	"lnVR_se", "lnRR", "lnRR_lower", "lnRR_upper" ,"lnRR_se")

>>>>

2) Subsetting the data to choose only one data point per individual per trait
```{r}
# this is a necessary step for the loop across all traits 
data_subset_parameterid_individual_by_age <- function(mydata, parameter, age_min, age_center) {
  tmp <- mydata %>%
    filter(age_in_days >= age_min,
           id == parameter) %>%
    # take results for single individual closest to age_center
    mutate(age_diff = abs(age_center - age_in_days)) %>%
    group_by(biological_sample_id) %>%
    filter(age_diff == min(age_diff)) %>%
    select(-age_diff)

@@

## Heterogeneity
#### FIGURE 4 (Second-order meta analysis on heterogeneity)
#### Create matrix to store results for all traits
```{r}
results.allhetero.grouping <- as.data.frame(cbind(c(1:n), matrix(rep(0, n*30), ncol = 30))) 
names(results.allhetero.grouping) <- c("id", "sigma2_strain.CVR", "sigma2_center.CVR", "sigma2_error.CVR", "s.nlevels.strain.CVR", 
	"s.nlevels.center.CVR", "s.nlevels.error.CVR", "sigma2_strain.VR", "sigma2_center.VR", "sigma2_error.VR", "s.nlevels.strain.VR", 
	"s.nlevels.center.VR", "s.nlevels.error.VR", "sigma2_strain.RR", "sigma2_center.RR", "sigma2_error.RR", "s.nlevels.strain.RR", 
	"s.nlevels.center.RR", "s.nlevels.error.RR", "lnCVR", "lnCVR_lower", "lnCVR_upper", "lnCVR_se", "lnVR", "lnVR_lower", "lnVR_upper", 
	"lnVR_se", "lnRR", "lnRR_lower", "lnRR_upper" ,"lnRR_se")
